---
title: "Chapter 2 Names and Values"
author: "Billy Caughey"
date: "8/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(1014)

library(lobstr)

```

# Chapter 2.1 Introduction

## Quiz 

### Question 1

_Given the following data frame, how do I create a new column called “3” that contains the sum of 1 and 2? You may only use $, not [[. What makes 1, 2, and 3 challenging as variable names?_

```{r}

df <- data.frame(runif(3), runif(3))
names(df) <- c(1, 2)

df$`3` <- df$`1` + df$`2`

print(df)

```

Using the variable names of 1, 2, and 3 are hard because variables should not start with numbers. Additionally, bracket notation becomes difficult.

### Question 2

_In the following code, how much memory does `y` occupy?_

```{r}
# old solution
x <- runif(1e6)

y <- list(x, x, x)

object.size(y)

#new solution

library(lobstr)
obj_size(y)

```

### Question 3 

_On which line does `a` get copied in the following example?_

Old answer: `a` is copied in the second line when it is copied into `b`.
New Answer: `a` is copied in the third line using the _copy on modify_ principle.

```{r}

a <- c(1, 5, 3, 2)
b <- a
b[[1]]

```

## 2.2 Binding Basics 

Consider the following:

```{r}

x <- c(1,2,3)
y <- x

obj_addr(x)
obj_addr(y)

```

### 2.2.1 Non-syntactic Names

Non-syntactic names are names that:

* Start with a '_' or digit
* A reserved word with the use of backticks (`)

I can get around the non-syntatic names by using the backtips. This is __not__ recommended. In fact, it's down right discouraged.

In fact, the following code chunks won't run in markdown (I'm not going to change the options):

* `_abc <- 1`: This is non-syntatic because it starts with a '_'.
* `if <- 10`: This is trying to bind the value 10 to the keyword __if__.

### 2.2.2 Exercises

#### Question 1

_Explain the relationship between `a`, `b`, `c`, and `d` in the following code:

```{r}

a <- 1:10
b <- a
c <- b
d <- 1:10

```

First, a numeric vector, `1:10` is bound to `a`. Then, `b` is bound to `1:10`. In the third line, `c` is bound to `1:10`. The final line sees the variable `d` bound to a different vector of `1:10`.

#### Question 2

_The following code accesses the mean function in multiple ways. Do they all point to the same underlying function object? Verify this with `lobstr::obj_addr()`_.

```{r}

obj_addr(mean)
obj_addr(base::mean)
obj_addr(get("mean"))
obj_addr(evalq(mean))
obj_addr(match.fun("mean"))

```

Yes, `obj_addr` calls the same __mean__ object.

#### Question 3

_By default, base R data import functions,like __read.csv()__, will automatically convert non-syntactic names to syntactic ones. Why might this be problematic. What option allows you to suppress this behaviour?

This is problematic because it could change the meaning of the name. Naming conventions matter. When naming conventions in a CSV are changed, meaning could be lost if a data dictionary is not found.

Within the `read.csv` function, there is an argument that allows the user to user their own field names. Additionally, it is entirely possible to rename the variable names once read in.

## 2.3 Copy-on-modify

Consider the following code:

```{r}

x <- c(1,2,3)
y <- x

obj_addr(x)
obj_addr(y)

y[[3]] <- 4
obj_addr(y)

```

Notice how x and y have the same address initially? This is because the line `x <- c(1,2,3)` creates a vector of 1, 2, and 3 then binds it x. The next line, `y <- x`, binds the same vector to y. Essentially, the vector of 1, 2, and 3 has TWO names it can be called by now. Once I make a modification, line `y[[3]] <- 4`, R creates a new vector of 1, 2, and 4 and binds that vector to y. This behavior is called __copy-on-modify__.

### 2.3.1 tracemem()

I can use the `tracemem()` function to see when an object is copied.

Consider the following code:

```{r}

x <- c(1, 2, 3)
cat(tracemem(x), "\n")

y <- x
y[[3]] <- 4

```

Notice how it states [location] -> [location]? This is telling me that R has made a copy and where it's new address is.

Also make sure to use the `untracemem()` function, or I'll keep getting the address calls.

```{r}

untracemem(x)

```

### 2.3.2 Function calls 

The same __copy-on-modify__ behavior exists in functions as well. Consider the following code:

```{r}

f <- function(a) {
    a
}

x <- c(1,2,3)
cat(tracemem(x), "\n")

z <- f(x)

untracemem(x)

```

There isn't a copy made in the line `z <- f(x)`, hence R doesn't suggest anything has happened.

### 2.3.3 Lists 

I shouldn't be naive enough to believe that only names (e.g., variables) point to values; elements of lists do to! Consider the following code:

```{r}

l1 <- list(1, 2, 3)
l2 <- l1
ref(l1, l2)

l2[[3]] <- 4
ref(l1, l2)

```

The schematic looks a bit interesting, but lets walk through it. First, when I bind the list of `list(1,2,3)` to l1 and l2, all the elements of those lists are in the same location. When I make a modification, look what happens. Only the third element of list `l2` changes address. This is because it has been modified.

### 2.3.4 Data frames 

Data frames are lists of vectors, so __copy-on-modify__ has important consequences when you modify a data frame. Consider the following code:

```{r}

d1 <- data.frame(x = c(1, 5, 6), y = c(2, 4, 3))
d2 <- d1
ref(d1, d2)

d2[, 2] <- d2[, 2] * 2
ref(d1, d2)

```

The same result that we saw in the list example is occuring in this example.

Now, the question is what happens if I modify every row instead of a single column? Let's find out:

```{r}

d3 <- d1
d3[1, ] <- d3[1, ] * 3

ref(d1, d3)

```

In this case, both fields in `d3` have been modified so they have new addresses.

### 2.3.5 Character vectors 

R uses references with character vectors as well. Consider the following:

```{r}

x <- c("a", "a", "abc", "d")
ref(x, character = TRUE)

```

It is important to note that R actually has a __global string pool__ that it binds variables to. That is why `"a"` has the same address for both elements. 

### 2.3.6 Exercises 

#### Question 1

_Why is `tracemem(1:10)` not useful?_

`tracemem(1:10)` isn't useful because each time I run this code, a different vector of 1, 2, 3, ... , 10 is created. That would mean new addresses each time I ran it.

```{r}

tracemem(1:10)
tracemem(1:10)

```

#### Question 2 

_Explain why tracemem() shows two copies when you run this code. Hint: carefully look at the difference between this code and the code shown earlier in this section._

Consider the code:

```{r}

x <- c(1L, 2L, 4L)
tracemem(x)

x[[3]] <- 4
untracemem(x)

```

The reason for the change is I'm changing from an integer of `4L` to a double in `4`.

#### Question 3

_Sketch out the relationship between the following objects._

```{r} 

a <- 1:10
b <- list(a, a)
c <- list(b, a, 1:10)

ref(a, b, c)

```

#### Question 4 

_What happens when you run this code? Draw a picture._

I am binding the second element of `x` to the list `x` is bound to.

```{r}

x <- list(1:10)
x[[2]] <- x

ref(x)

x

```

## 2.4 Object Size 

I can use the function `obj_size()` to find out how much memory an object takes.

```{r}

obj_size(letters)
obj_size(ggplot2::diamonds)

```

Since lists are references to values, the size of the list might be smaller than I think.

```{r}

x <- runif(1e6)
obj_size(x)

y <- list(x, x, x)
obj_size(y)

```

With characters, since R uses the global string pool, the character vectors take up less memory than I think.

```{r}

banana <- "bananas bananas bananas"
obj_size(banana)

obj_size(rep(banana, 100))

```

One word of caution: `obj_size(x) + obj_size(y) = obj_size(x, y)` __IF THEY DON'T HAVE COMMON VALUES__.

```{r}

obj_size(x, y)

x <- c(1, 2, 4)
obj_size(x)
y <- c(5, 6, 7)
obj_size(y)
obj_size(x, y)

```

### 2.4.1 Exercises 

#### Question 1 

_In the following example, why are `object.size()` and `obj_size()` so radically different? Consult the documentation of `object.size()`._

```{r}

y <- rep(list(runif(1e4)), 100)

object.size(y)
obj_size(y)

```

`obj_size` is considering the memory allocation for y and all the nuances of ALTREP. `object.size` brute forces the size without nuances.

#### Question 2

_Take the following list. Why is its size somewhat misleading?_

```{r}

funs <- list(mean, sd, var)
obj_size(funs)

obj_size(mean)
obj_size(sd)
obj_size(var)

```

It's misleading because the `obj_size` isn't the sum of it's parts. These are three functions, but `sd` and `var` call `mean`. With those calls, the actual memory allocation needed is smaller than the some of it's parts.

#### Question 3

_Predict the output of the following code:_

```{r}

a <- runif(1e6)
obj_size(a)

b <- list(a, a)
obj_size(b)
obj_size(a, b)

b[[1]][[1]] <- 10
obj_size(b)
obj_size(a, b)

b[[2]][[1]] <- 10
obj_size(b)
obj_size(a, b)

```

## 2.5 Modify-in-place

There are two exceptions to `modify-in-place`:

1. Objects with a single binding get a special performance optimization
2. Environments, a special type of object, are always modified in place

### 2.5.1 Objects with a single binding

If an object has a single name bound to it, R will modify it in place:

```{r}

v <- c(1, 2, 3)
cat(tracemem(v), "\n")

v[[3]] <- 4
untracemem(v)

```

There are two complications:

1. When it comes to binding, R can currently only count 0, 1, and many. This means if there are two bindings, and one goes away, there are still many because many - 1 is many. this means R will make copies when it sometimes doesn't need to.

2. Whenever you call the vast majority of functions, it makes a reference to the object.

These subtleties can be explored with using a case study:

```{r}

x <- data.frame(matrix(runif(5 * 1e4), ncol = 5))
medians <- vapply(x, median, numeric(1))

cat(tracemem(x), "\n")

for(i in seq_along(medians)){
    x[[i]] <- x[[i]] - medians[[i]]
}

untracemem(x)

```

The number of copies can be reduced by using a `list` instead of a `data.frame`.

```{r}

y <- as.list(x)
cat(tracemem(y),"\n")

for(i in 1:5){
    y[[i]] <- y[[i]] - medians[[i]]
}

untracemem(y)

```




















































